\begin{longtable}{p{3cm}p{2cm}p{5.7cm}p{2.5cm}}
    \caption{Datasets used in the experiments} \\
    \label{tbl:datasets} \\
    \toprule
    \textbf{Dataset}   & \textbf{Type} & \textbf{Description} & \textbf{\#Features} \\
    \midrule
    \endfirsthead

    \multicolumn{4}{c}{{\bfseries Table \thetable\ continued from previous page}} \\
    \toprule
    \textbf{Dataset}   & \textbf{Type} & \textbf{Description} & \textbf{\#Features} \\
    \midrule
    \endhead

    \midrule
    \multicolumn{4}{r}{Continued on next page} \\
    \midrule
    \endfoot

    \bottomrule
    \endlastfoot

    Buildings          & Polygon       & Contains building outlines. Its \emph{type} attribute can have values like \emph{house}, \emph{university}, and \emph{restaurant}. & 4,147,645 \\
    Land Use           & Polygon       & Represents areas designated to different purposes and activities. Its \emph{fclass} attribute can have values like \enquote{forest}, \enquote{farmland}, and \enquote{residential}. & 541,452 \\
    Natural            & Point         & Contains outlines of various objects found in nature. Its \emph{fclass} attribute can have values like \enquote{beach}, \enquote{glacier}, and \enquote{cave\_entrance}. & 119,725 \\
    Natural            & Polygon       & Similar to the point data equivalent. & 6,665 \\
    Places of Worship  & Point         & Common values for \emph{fclass} attribute: \enquote{christian}, \enquote{buddhist}, and \enquote{muslim}. & 311 \\
    Places of Worship  & Polygon       & Similar to the point data equivalent. & 2,520 \\
    Places             & Point         & Common values for \emph{fclass} attribute: \enquote{farm}, \enquote{village}, and \enquote{island}. Repeated entries trimmed for brevity. & 178,997 \\
    Places             & Polygon       & Similar to the point data equivalent. & 5,689 \\
    Points of Interest & Point         & Common values for \emph{fclass} attribute: \enquote{tourist\_info}, \enquote{bench}, and \enquote{kindergarten}. & 117,677 \\
    Points of Interest & Polygon       & Similar to the point data equivalent. & 45,825 \\
    Railways           & Lines         & Common values for \emph{fclass} attribute: \enquote{rail}, \enquote{subway}, and \enquote{tram}. Also has True/False attributes \emph{bridge} and \emph{tunnel}. & 14,008 \\
    Roads              & Lines         & Common values for \emph{fclass} attribute: \enquote{rail}, \enquote{subway}, and \enquote{tram}. Has additional attributes \emph{oneway}, \emph{maxspeed}, \emph{bridge}, and \emph{tunnel}. & 1,741,929 \\
    Traffic            & Point         & Common values for \emph{fclass} attribute: \enquote{crossing}, \enquote{street\_lamp}, and \enquote{parking}. & 98,860 \\
    Traffic            & Polygon       & Common values for \emph{fclass} attribute: \enquote{parking}, \enquote{pier}, and \enquote{dam}. & 45,623 \\
    Transport          & Point         & Common values for \emph{fclass} attribute: \enquote{bus\_stop}, \enquote{ferry\_terminal}, and \enquote{railway\_station}. & 91,627 \\
    Transport          & Polygon       & Similar to the point data equivalent. & 935 \\
    Water              & Polygon       & Common values for \emph{fclass} attribute: \enquote{water}, \enquote{wetland}, and \enquote{river\_bank}. & 1,861,199 \\
    Waterways          & Lines         & Common values for \emph{fclass} attribute: \enquote{stream}, \enquote{river}, and \enquote{canal}. & 833,253 \\
\end{longtable}
