\section*{Preface}

\begin{comment}
The Preface includes the facts: what type of project, where it is conducted,
who supervised, and any acknowledgements you wish to give.

This Master's Thesis template was created by Bj\"orn Gamb\"ack and is based on a template that he created for the 2016 ``Experts in Team'' course on
Computational Creativity (TDT4853) at the Norwegian University of Science and Technology (NTNU),
which in turn was heavily based on the 2014 AI Master's Thesis template created by Anders Kofod-Petersen ---
with some of the explaining text stemming from Anders' original template.

You may basically thank anybody you like (and avoid thanking anybody you do not like) and in any form you like.
However, it is a good idea to always thank people who made direct contributions, e.g., those whose data you have been given access to or those whose images you have been given permission to reproduce.

Some students choose to include the text of the original project description in the Preface. This is possible but not necessary,
in particular not if you have changed the theme somewhat over time.
The Preface of the Master's Thesis might also be a good place to introduce your Specialisation Project, in case you plan
on reusing some texts from it (since the Specialisation Project is not a published and easily accessible work, and might
not be known to your audience, neither your text in itself nor even the general concept as such).
\end{comment}

This master's thesis is written at the Norwegian University of Science and Technology, for the Division of Geomatics at the Department of Civil and Environmental Engineering. It is the final work of the study programme called Engineering and ICT.

I would like to thank Hongchao Fan, my supervisor at the Division of Geomatics, and my two co-supervisors from Norkart: Alexander Salveson Nossum and Arild Nomeland. I would also like to thank GeoForum and Digin for inviting me to present my thesis at Geomatikkdagene and GeoAI:Konferansen, respectively. Finally, I would like to thank my ever-supporting family for their constant encouragement and support throughout my studies.

\vfill

\hfill \thesisAuthor

\hfill Trondheim, \today
